---
title: "Turnstiles"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Packages

```{r packages}
library(tidyverse)
library(tidycensus)
library(sf)
library(lubridate)
library(gganimate)
```

## Importing Data
https://medium.com/qri-io/taming-the-mtas-unruly-turnstile-data-c945f5f96ba0 was an important resource for this section.

http://web.mta.info/developers/developer-data-terms.html#data provided much of the data.

```{r message=FALSE, warning=FALSE}
# turnstile_data <- read_delim("turnstile_200201.txt", ",") 
# start <- as.Date("02-08-20",format="%m-%d-%y")
# end   <- as.Date("05-23-20",format="%m-%d-%y")
# theDate <- start
# while (theDate <= end) {
#   temp =read_delim(paste0("turnstile_", format(theDate,"%y%m%d"), ".txt"), ",")
#   turnstile_data <- turnstile_data %>% bind_rows(temp)
#   print(theDate)
#   theDate <- theDate + 7                   
# }
# 
# turnstile_data <- turnstile_data %>%
#   mutate(DATE = mdy(DATE),
#          ENTRIES = as.numeric(ENTRIES)) %>% rename(EXITS = `EXITS                                                               `) %>% mutate(EXITS = as.numeric(EXITS)) %>%
#   mutate(DATETIME = ymd_hms(paste(DATE, TIME)))
# turnstile_data <- turnstile_data %>%
#   mutate(id = paste(`C/A`, UNIT, SCP, DATE, TIME)) %>%
#   mutate(unit_id = paste(`C/A`, UNIT, SCP)) %>%
#   arrange(id)
```

```{r}
load("turnstile_data")
```


```{r}
turnstile_data <- turnstile_data %>%
  group_by(unit_id) %>%
  mutate(net_entries = abs(ENTRIES - lag(ENTRIES, 1)),
         net_exits = abs(EXITS - lag(EXITS, 1))) %>%
  mutate(net_entries = replace(net_entries, net_entries > 10000, 0)) %>% 
  mutate(net_exits = replace(net_exits, net_exits > 10000, 0)) %>% 
  ungroup()
  # group_by(unit_id) %>%
  # mutate(net_entries = abs(lead(ENTRIES, 1) - ENTRIES),
  #        net_exits = abs(lead(EXITS, 1) - EXITS)) %>%
  # mutate(net_entries = replace(net_entries, net_entries > 10000, 0)) %>% 
  # mutate(net_exits = replace(net_exits, net_exits > 10000, 0)) %>% 
  # ungroup()
```


Credits to Chris Whong for compiling the remote_complex_lookup data table.
```{r}
station_booths <- readxl::read_excel("Remote-Booth-Station.xls")
remote_complex_lookup <- read_csv("remote_complex_lookup.csv")
turnstile_data_joined <- turnstile_data %>% left_join(remote_complex_lookup,
                                                       by = c("UNIT" = "remote",
                                                              "C/A" = "booth"))
stations <- read_csv("Stations.csv")
turnstile_data_joined_stations <-
  turnstile_data_joined %>%
  left_join(
    stations %>%
      group_by(`Complex ID`) %>%
      filter(row_number() == 1) %>%
      ungroup(),
    by = c("complex_id" = "Complex ID")
  ) %>%
  filter(DIVISION != "PTH", DIVISION != "RIT")

turnstile_summaries <-
  turnstile_data_joined_stations %>%
  group_by(round_date(DATETIME, "4 hours")) %>%
  filter(n() > 5) %>% summarize(entry_sum = sum(net_entries, na.rm = TRUE),
                                exit_sum = sum(net_exits, na.rm = TRUE)) %>% 
  rename(datetime = 1)
```


#### Figure 1: Total Daily Turnstile Entries and Exits
```{r}
ggplot(
  data = turnstile_summaries %>% group_by(the_date = date(datetime)) %>% summarize(
    entry_sum = sum(entry_sum),
    exit_sum = sum(exit_sum)
  ) %>% pivot_longer(ends_with("_sum"), names_to = 'sumtype', values_to = 'the_sum')
) +
  geom_line(mapping = aes(
    x = the_date,
    y = the_sum,
    group = sumtype,
    color = sumtype
  )) +
  geom_point(mapping = aes(
    x = the_date,
    y = the_sum,
    group = sumtype,
    color = sumtype
  )) +
  labs(
    title = "Total Daily Turnstile Entries and Exits",
    x = "Date",
    y = "Total Daily Count",
    color = "Legend",
    subtitle = paste(
      "Orange Vertical Line is when State of Emergency was Declared in NYC (March 12)",
      "Red Vertical Line is when Statewide Stay-At-Home Order was Enacted (March 22)",
      sep = "\n"
    )
  ) +
  scale_color_manual(values = c("#7cae00", "#00bfc4"), labels = c("Entries", "Exits")) +
  theme(legend.position = "right") +
  scale_y_continuous(labels = scales::comma) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-12-2020")),
    color = "dark orange",
    linetype = "dashed"
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-22-2020")),
    color = "red",
    linetype = "dashed"
  )
```


```{r}
entries_transition <- ggplot(data = turnstile_summaries) + 
  geom_line(mapping = aes(x = datetime, y = entry_sum)) +
  geom_point(mapping = aes(x = datetime, y = entry_sum)) +
  transition_states(
    hour(datetime),
    transition_length = 0,
    state_length = 1
  ) +
  labs(title = 'Total Entries into Subway System by Time of Day', subtitle = 'Hour of Day: {closest_state}', x = 'Date', y = 'Total System Entries')

exits_transition <- ggplot(data = turnstile_summaries) + 
  geom_line(mapping = aes(x = datetime, y = exit_sum)) +
  geom_point(mapping = aes(x = datetime, y = exit_sum)) +
  transition_states(
    hour(datetime),
    transition_length = 0,
    state_length = 1
  ) +
  labs(title = 'Total Exits from Subway System by Time of Day', subtitle = 'Hour of Day: {closest_state}', x = 'Date', y = 'Total System Exits')
```

```{r}
# animate(entries_transition, height = 5, width = 5, units = "in", res = 150)
# anim_save("entries_transition.gif")
# animate(exits_transition, height = 5, width = 5, units = "in", res = 150)
# anim_save("exits_transition.gif")
```

```{r message=FALSE, warning=FALSE}
ggplot(data = turnstile_summaries %>%
         pivot_longer(ends_with("_sum"), names_to = 'sumtype', values_to = 'the_sum')) +
  geom_line(mapping = aes(
    x = datetime,
    y = the_sum,
    group = sumtype,
    color = sumtype
  )) +
  geom_point(mapping = aes(
    x = datetime,
    y = the_sum,
    group = sumtype,
    color = sumtype
  )) +
  scale_y_continuous(labels = scales::comma) +
facet_wrap(vars(hour(datetime))) +
  labs(
    title = 'Total Turnstile Entries and Exits by Time of Day',
    x = 'Date',
    y = 'Total Count',
    color = "Legend",
    subtitle = paste(
      "Orange Vertical Line is when State of Emergency was Declared in NYC (March 12)",
      "Red Vertical Line is when Statewide Stay-At-Home Order was Enacted (March 22)",
      sep = "\n"
    )
  ) +
  scale_color_manual(values = c("#7cae00", "#00bfc4"),
                     labels = c("Entries", "Exits")) +
  theme(legend.position = "right") +
  geom_vline(
    mapping = aes(xintercept = mdy("03-12-2020") %>% as_datetime()),
    color = "dark orange",
    linetype = "dashed"
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-22-2020") %>% as_datetime()),
    color = "red",
    linetype = "dashed"
  )
```

## Hourly Summaries and Rates
```{r}
# get summary by hour for the entirety of the selected period
get_hourly_summaries <- function(periodtype, filterperiod, filterhour, summaryobject) {
  if (periodtype == "date") {
    temp = date(summaryobject$datetime)
  } else if (periodtype == "week") {
    temp = week(summaryobject$datetime)
  } else if (periodtype == "month") {
    temp = month(summaryobject$datetime)
  }
  time_filtered_summaryobject <- summaryobject %>% filter(
    hour(datetime) == filterhour,
    temp == filterperiod)
  weekday_only_summaryobject <-
    time_filtered_summaryobject %>% filter(!(wday(datetime, label = TRUE) %in% c("Sat", "Sun")))
  weekend_only_summaryobject <-
    time_filtered_summaryobject %>% filter((wday(datetime, label = TRUE) %in% c("Sat", "Sun")))
  if (nrow(weekday_only_summaryobject) != 0) {
    weekday_hourly_entry_avg = weekday_only_summaryobject$entry_sum %>% mean(na.rm = TRUE)
    weekday_hourly_exit_avg = weekday_only_summaryobject$exit_sum %>% mean(na.rm = TRUE)
  } else {
    weekday_hourly_entry_avg = NA
    weekday_hourly_exit_avg = NA
  }
  if (nrow(weekend_only_summaryobject) != 0) {
    weekend_hourly_entry_avg = weekend_only_summaryobject$entry_sum %>% mean(na.rm = TRUE)
    weekend_hourly_exit_avg = weekend_only_summaryobject$exit_sum %>% mean(na.rm = TRUE)
  } else {
    weekend_hourly_entry_avg = NA
    weekend_hourly_exit_avg = NA
  }
  entry_avgs <- c(weekday_hourly_entry_avg, weekend_hourly_entry_avg)
  exit_avgs <- c(weekday_hourly_exit_avg, weekend_hourly_exit_avg)
  toReturn = tibble(
      c("Weekday", "Weekend"),
      c(filterperiod, filterperiod),
      c(filterhour, filterhour),
      entry_avgs,
      exit_avgs,
      c(periodtype, periodtype)
  ) %>% rename(DayType = 1, Period = 2, Hour = 3, EntryAvgs = 4, ExitAvgs = 5, PeriodType = 6)
  return(toReturn)
}

get_percentage_of_period <- function(summaryobject, periodtype) {
  if (periodtype == "month") {
    periods <- unique(month(summaryobject$datetime))
  } else if (periodtype == "week") {
    periods <- unique(week(summaryobject$datetime))
  } else if (periodtype == "date") {
    periods <- unique(date(summaryobject$datetime))
  }
  hours <- c(0,4,8,12,16,20)
  febavg <- NULL
  for (currhour in hours) {
    febavg <- febavg %>% 
      bind_rows(get_hourly_summaries("month", 2, currhour, summaryobject))
  }
  febavg <-
    febavg %>% group_by(DayType) %>% mutate(
      EntryAvgsTotal = sum(EntryAvgs),
      ExitAvgsTotal = sum(ExitAvgs),
      ProportionOfDailyEntry = EntryAvgs / EntryAvgsTotal,
      ProportionOfDailyExit = ExitAvgs / ExitAvgsTotal
    ) %>%
    ungroup() %>% 
    rename(FebEntryAvgs = EntryAvgs, 
           FebExitAvgs = ExitAvgs,
           FebProportionOfDailyEntry = ProportionOfDailyEntry,
           FebProportionOfDailyExit = ProportionOfDailyExit) %>% 
    select(-c(EntryAvgsTotal, ExitAvgsTotal, Period, PeriodType))
  toReturn <- NULL
  for (currperiod in periods) {
    for (currhour in hours) {
      toReturn <-
        toReturn %>% bind_rows(get_hourly_summaries(periodtype, currperiod, currhour, summaryobject))
    }
  }
  toReturn <-
    toReturn %>% group_by(DayType, Period) %>% mutate(
      EntryAvgsTotal = sum(EntryAvgs),
      ExitAvgsTotal = sum(ExitAvgs),
      ProportionOfDailyEntry = EntryAvgs / EntryAvgsTotal,
      ProportionOfDailyExit = ExitAvgs / ExitAvgsTotal
    ) %>%
    ungroup() %>% 
    left_join(febavg, by=c("DayType", "Hour")) %>%
    drop_na(EntryAvgs) %>% mutate(
      EntryAvgsChange = FebEntryAvgs - EntryAvgs,
      ExitAvgsChange = FebExitAvgs - ExitAvgs,
      PercentageOfDailyEntryChange = 100*(FebProportionOfDailyEntry - ProportionOfDailyEntry),
      PercentageOfDailyExitChange = 100*(FebProportionOfDailyExit - ProportionOfDailyExit)
    )
  return(toReturn)
}

# as_date(18293) ==> "2020-02-01"
# NOTE: period is always stored as an integer
turnstile_date_percentages <- get_percentage_of_period(turnstile_summaries, "date")
turnstile_week_percentages <- get_percentage_of_period(turnstile_summaries, "week")
turnstile_month_percentages <- get_percentage_of_period(turnstile_summaries, "month")
```
EntryRatioToEntireDayAvg gives the ratio between average weekday/weekend entries for that hour and the total average weekday/weekend entries



```{r}
ggplot(data = turnstile_week_percentages) +
  geom_point(aes(
    x = as_date("2020-01-01") + (7*turnstile_week_percentages$Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour)
  )) +
  geom_line(aes(
    x = as_date("2020-01-01") + (7*turnstile_week_percentages$Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour),
    group = as.factor(Hour)
  )) +
  facet_wrap(vars(DayType)) +
  labs(
    title = "Proportion of Daily Total Turnstile Entries by Time Interval",
    x = "Month",
    y = "Proportion of All Rides",
    color = "Interval Ending At",
    subtitle = paste(
      "Orange Vertical Line is when State of Emergency was Declared in NYC (March 12)",
      "Red Vertical Line is when Statewide Stay-At-Home Order was Enacted (March 22)",
      sep = "\n"
    )
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-12-2020")),
    color = "dark orange",
    linetype = "dashed"
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-22-2020")),
    color = "red",
    linetype = "dashed"
  )
  
```

```{r}
ggplot(data = turnstile_date_percentages) +
  geom_point(aes(
    x = as_date(Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour)
  )) +
  geom_line(aes(
    x = as_date(Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour),
    group = as.factor(Hour)
  )) +
  facet_wrap(vars(DayType)) +
  labs(
    title = "Proportion of Daily Total Turnstile Entries by Time Interval (Daily Averages)",
    x = "Month",
    y = "Proportion of All Rides",
    color = "Interval Ending At",
    subtitle = paste(
      "Orange Vertical Line is when State of Emergency was Declared in NYC (March 12)",
      "Red Vertical Line is when Statewide Stay-At-Home Order was Enacted (March 22)",
      sep = "\n"
    )
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-12-2020")),
    color = "dark orange",
    linetype = "dashed"
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-22-2020")),
    color = "red",
    linetype = "dashed"
  )
```


```{r}
ggplot(data = turnstile_month_percentages %>% filter(Period %in% c(2, 4)) %>% 
  mutate(Period = month.abb[Period])) +
  geom_point(aes(x = ((Hour - 2) + 24) %% 24,
                 y = ProportionOfDailyEntry,
                 color = as.factor(Period))) +
  geom_line(aes(x = ((Hour - 2) + 24) %% 24,
                y = ProportionOfDailyEntry,
                color = Period,
                group = as.factor(Period))) +
  facet_grid(. ~ DayType) +
  labs(
    x = "Time Of Day",
    y = "Proportion of All Rides",
    title = "Proportion of Daily Total Turnstile Entries by Time Interval in February and April",
    subtitle = "Points are at Time Interval Midpoints for Easier Visual Interpretation",
    color = "Month",
    caption = "For instance, 10am in this chart corresponds to 12pm in the previous chart"
  )
```

```{r}
view(turnstile_month_percentages %>% filter(Period == 2))
```





## Incorporating Census Data
```{r}
# curr_vars <- load_variables(2018, "acs5")
bronxzip <- as.character(c(10453, 10457, 10460, 10458, 10467, 10468, 10451, 10452, 10456, 10454, 10455, 10459, 10474, 10463, 10471, 10466, 10469, 10470, 10475, 10461, 10462, 10464, 10465, 10472, 10473))
brooklynzip <- as.character(c(11212, 11213, 11216, 11233, 11238, 11209, 11214, 11228, 11204, 11218, 11219, 11230, 11234, 11236, 11239, 11223, 11224, 11229, 11235, 11201, 11205, 11215, 11217, 11231, 11203, 11210, 11225, 11226, 11207, 11208, 11211, 11222, 11220, 11232, 11206, 11221, 11237))
manhattanzip <- as.character(c(10026, 10027, 10030, 10037, 10039, 10001, 10011, 10018, 10019, 10020, 10036, 10029, 10035, 10010, 10016, 10017, 10022, 10012, 10013, 10014, 10004, 10005, 10006, 10007, 10038, 10280, 10002, 10003, 10009, 10021, 10028, 10044, 10065, 10075, 10128, 10023, 10024, 10025, 10031, 10032, 10033, 10034, 10040))
queenszip <- as.character(c(11361, 11362, 11363, 11364, 11354, 11355, 11356, 11357, 11358, 11359, 11360, 11365, 11366, 11367, 11412, 11423, 11432, 11433, 11434, 11435, 11436, 11101, 11102, 11103, 11104, 11105, 11106, 11374, 11375, 11379, 11385, 11691, 11692, 11693, 11694, 11695, 11697, 11004, 11005, 11411, 11413, 11422, 11426, 11427, 11428, 11429, 11414, 11415, 11416, 11417, 11418, 11419, 11420, 11421, 11368, 11369, 11370, 11372, 11373, 11377, 11378))
sizip <- as.character(c(10302, 10303, 10310, 10306, 10307, 10308, 10309, 10312, 10301, 10304, 10305, 10314))
nyczip <- as.character(c(bronxzip, brooklynzip, manhattanzip, queenszip, sizip))

zipcode_data <- get_acs(
  geography = "zip code tabulation area",
  variables = c(MedianIncome = "B19013_001",
                TotalPopulation = "B01003_001",
                TotalHousingUnits = "B25001_001",
                TransitCommuteNum = "B08301_010",
                AggregateTransportTime = "B08136_001"),
  year = 2018
) %>% filter(GEOID %in% nyczip) 

zipcode_data_spread <- zipcode_data %>%
  pivot_wider(
    id_cols = c(GEOID, NAME),
    names_from = variable,
    values_from = c(estimate)
  )

zipcode_shapefiles <- get_acs(
  geography = "zip code tabulation area",
  variables = c(TotalPopulation = "B01003_001"),
  geometry = TRUE,
  keep_geo_vars = TRUE,
  year = 2018
) %>% filter(GEOID %in% nyczip) %>% select(-c(estimate, moe, variable))

zipcode_data_joined <- zipcode_data_spread %>% inner_join(zipcode_shapefiles) %>% 
  mutate(PopDensity = TotalPopulation / (ALAND10 / 2589988),
         TransitUsageRate = TransitCommuteNum / TotalPopulation,
         AvgTransportTime = AggregateTransportTime / TotalPopulation)
zipcode_data_sf <-zipcode_data_joined %>% st_sf() %>% st_transform(4326)
zipcode_data_transformed <- zipcode_data_sf %>%  
  st_transform(2831)
```

```{r}
turnstile_data_grouped <- turnstile_data_joined_stations %>% 
  group_by(complex_id, DATETIME) %>%
  summarize(latitude = mean(`GTFS Latitude`), 
            longitude = mean(`GTFS Longitude`),
            net_entries = sum(net_entries, na.rm = TRUE),
            net_exits = sum(net_exits, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(!is.na(latitude))

turnstile_data_sf_together <-
  turnstile_data_grouped %>% 
  group_by(complex_id) %>% 
  summarize(latitude = mean(`latitude`), 
            longitude = mean(`longitude`),
            net_entries = sum(net_entries, na.rm = TRUE),
            net_exits = sum(net_exits), na.rm = TRUE) %>% 
  filter(!is.na(latitude)) %>% 
  st_as_sf(coords = c("longitude", "latitude"))
st_crs(turnstile_data_sf_together) <- 4326
turnstile_data_sf_together_transformed <-
  turnstile_data_sf_together %>% st_transform(2831)
  
```

```{r}
combined_table <-
  st_join(turnstile_data_sf_together, zipcode_data_sf)
combined_table_nas <-
  combined_table %>% filter(is.na(GEOID)) %>% select(c(1:7)) %>%
  st_join(zipcode_data_sf, join = st_nearest_feature)
combined_table <-
  combined_table %>% drop_na(GEOID) %>%  bind_rows(combined_table_nas)
```


### Actual Zip-Code Joining For Paper

```{r}
turnstile_data_grouped <- turnstile_data_joined_stations %>% 
  group_by(complex_id, DATETIME) %>%
  summarize(latitude = mean(`GTFS Latitude`), 
            longitude = mean(`GTFS Longitude`),
            net_entries = sum(net_entries, na.rm = TRUE),
            net_exits = sum(net_exits, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(!is.na(latitude))

turnstile_data_grouped_febonly <-
  turnstile_data_grouped %>% filter(month(DATETIME) == 2) %>%
  mutate(IsWeekend = wday(DATETIME, label = TRUE) %in% c("Sat", "Sun")) %>%
  mutate(DayType = if_else(IsWeekend, "Weekend", "Weekday")) %>%
  select(-IsWeekend)
daily_turnstile_data_summarized_febonly <-
  turnstile_data_grouped_febonly %>%
  mutate(datetime_round = round_date(DATETIME, "4 hours"), DayType) %>%
  # group_by(Hour = hour(datetime_round), DayType, complex_id) %>% 
  group_by(DayType, complex_id) %>% 
  summarize(
    Feb_net_entries = mean(net_entries, na.rm = TRUE),
    Feb_net_exits = mean(net_exits, na.rm = TRUE)
  ) %>% ungroup()

turnstile_data_grouped <-
  turnstile_data_grouped %>%
  mutate(IsWeekend = wday(DATETIME, label = TRUE) %in% c("Sat", "Sun")) %>%
  mutate(DayType = if_else(IsWeekend, "Weekend", "Weekday")) %>%
  select(-IsWeekend) %>%
  mutate(datetime_round = round_date(DATETIME, "4 hours"), DayType) %>%
  mutate(Hour = hour(datetime_round))
```

```{r}
daily_turnstile_data_grouped_summarized <- 
  turnstile_data_grouped %>% 
  group_by(Month = month(DATETIME), DayType, complex_id ) %>% #, Hour) %>% 
  summarize(net_entries = mean(net_entries, na.rm = TRUE),
            net_exits = mean(net_exits, na.rm = TRUE)) %>%
  inner_join(daily_turnstile_data_summarized_febonly) %>% 
  mutate(
    Diff_net_entries = net_entries - Feb_net_entries,
    Diff_net_exits = net_exits - Feb_net_exits,
    Diff_net_entries_percentage = 100 * (Diff_net_entries / Feb_net_entries),
    Diff_net_exits_percentage = 100 * (Diff_net_exits / Feb_net_exits)
  )
daily_turnstile_data_grouped_summarized_sf <- daily_turnstile_data_grouped_summarized %>% left_join(
    stations %>%
      group_by(`Complex ID`) %>%
      filter(row_number() == 1) %>%
      ungroup(),
    by = c("complex_id" = "Complex ID")
  ) %>% rename(latitude = `GTFS Latitude`,
               longitude = `GTFS Longitude`) %>% 
  filter(!is.na(latitude)) %>% 
  select(-c(`North Direction Label`, `South Direction Label`, Structure, `Daytime Routes`)) %>% 
  st_as_sf(coords = c("longitude", "latitude"))
st_crs(daily_turnstile_data_grouped_summarized_sf) <- 4326
  
  
daily_combined_table <-
  st_join(daily_turnstile_data_grouped_summarized_sf, zipcode_data_sf)
daily_combined_table_nas <-
  daily_combined_table %>% filter(is.na(GEOID)) %>% select(c(1:7)) %>%
  st_join(zipcode_data_sf, join = st_nearest_feature)
daily_combined_table <-
  daily_combined_table %>% drop_na(GEOID) %>%  bind_rows(daily_combined_table_nas)
daily_combined_table_sf <- daily_combined_table %>% st_as_sf()
st_crs(daily_combined_table_sf) <- 4326

station_complex_to_zip <- daily_combined_table %>% group_by(complex_id) %>% 
  summarize(Zip = first(GEOID), geometry = first(geometry), MedianIncome = first(MedianIncome),
            PopDensity = first(PopDensity)) %>% ungroup()
```

```{r warning=FALSE}
ggplot(
  data = daily_combined_table %>% filter(
    net_entries > 10,
    Feb_net_entries > 10,
    Month == 4,
    # Diff_net_entries_percentage < 0,
    DayType == "Weekday"
  ) %>% mutate(
    Borough = recode(
      Borough,
      Bx = "Bronx",
      Bk = "Brooklyn",
      M = "Manhattan",
      Q = "Queens",
      SI = "Staten Island"
    )
  )
) +
  geom_point(aes(x = `MedianIncome`, y = Diff_net_entries_percentage, color = Borough)) +
  labs(title = "Percentage Decline in Daily System Entries (Weekday) between February & April",
       subtitle = "Percentage change, by station",
       x = "Station Zip Code Median Income",
       y = "Percentage Change")
```

```{r message=FALSE, warning=FALSE}
ggplot(
  data = daily_combined_table %>% filter(
    net_entries > 10,
    Feb_net_entries > 10,
    Month == 4,
    # Diff_net_entries_percentage < 0,
    DayType == "Weekday"
  ) %>% mutate(
    Borough = recode(
      Borough,
      Bx = "Bronx",
      Bk = "Brooklyn",
      M = "Manhattan",
      Q = "Queens",
      SI = "Staten Island"
    )
  )
) +
  geom_point(aes(x = log(`MedianIncome`, 10), y = Diff_net_entries_percentage, color = Borough)) +
  labs(title = "Percentage Decline in Daily System Entries (Weekday) between February & April",
       subtitle = "Percentage change, by station",
       x = "log10(Station Zip Code Median Income)",
       y = "Percentage Change") + 
  geom_smooth(method='lm', aes(x = log(MedianIncome, 10), y = Diff_net_entries_percentage),
              se = FALSE, color = "black")
```



```{r}
ggplot(
  data = daily_combined_table %>% filter(
    net_entries > 10,
    Feb_net_entries > 10,
    Month == 4,
    Diff_net_entries_percentage < -25,
    DayType == "Weekend"
  ) %>% mutate(
    Borough = recode(
      Borough,
      Bx = "Bronx",
      Bk = "Brooklyn",
      M = "Manhattan",
      Q = "Queens",
      SI = "Staten Island"
    )
  )
) +
  geom_point(aes(x = `MedianIncome`, y = Diff_net_entries_percentage, color = Borough)) +
  labs(title = "Percentage Decline in Daily System Entries (Weekend) between February and April",
       subtitle = "Percentage change, by station",
       x = "Station Zip Code Median Income",
       y = "Percentage Change")
```

```{r}
ggplot(
  data = daily_combined_table_sf %>% filter(
    # net_entries > 10,
    # Feb_net_entries > 10,
    Month == 4,!is.na(Diff_net_entries_percentage),
    DayType == "Weekday"
  ) %>%  mutate(
    Borough = recode(
      Borough,
      Bx = "Bronx",
      Bk = "Brooklyn",
      M = "Manhattan",
      Q = "Queens",
      SI = "Staten Island"
    )
  )
) +
  labs(title = "Percentage Decline in Daily System Entries (Weekday) between February and April",
       subtitle = "Percentage change, by station",
       fill = "Median Zip Code Income ($)",
       color = "Percentage Change from February",
       size = "Average Daily Entries in February") +
  geom_sf(
    data = zipcode_data_sf,
    aes(fill = MedianIncome, geometry = geometry),
    size = 0.25,
    colour = "gray"
  ) +
  scale_fill_gradient(
    low = "#C9E3F2",
    high = "#071469",
    limits = c(0, 100000),
    oob = scales::squish
  ) +
  scale_color_gradient(
    low = "#8C0003",
    high = "pink",
    oob = scales::squish
  ) +
  geom_sf(
    aes(
      geometry = geometry,
      color = Diff_net_entries_percentage,
      size = Feb_net_entries
    ),
    shape = 16
  ) 
```
```{r}
ggplot(
  data = daily_combined_table_sf %>% filter(
    # net_entries > 10,
    # Feb_net_entries > 10,
    Month == 4,!is.na(Diff_net_entries_percentage),
    DayType == "Weekend"
  ) %>%  mutate(
    Borough = recode(
      Borough,
      Bx = "Bronx",
      Bk = "Brooklyn",
      M = "Manhattan",
      Q = "Queens",
      SI = "Staten Island"
    )
  )
) +
  labs(title = "Percentage Decline in Daily System Entries (Weekend) between February and April",
       subtitle = "Percentage change, by station",
       fill = "Median Zip Code Income ($)",
       color = "Percentage Change from February",
       size = "Average Daily Entries in February") +
  geom_sf(
    data = zipcode_data_sf,
    aes(fill = MedianIncome, geometry = geometry),
    size = 0.25,
    colour = "gray"
  ) +
  scale_fill_gradient(
    low = "#C9E3F2",
    high = "#071469",
    limits = c(0, 100000),
    oob = scales::squish
  ) +
  scale_color_gradient(
    low = "#8C0003",
    high = "pink",
    limits = c(-100, -60),
    oob = scales::squish
  ) +
  geom_sf(
    aes(
      geometry = geometry,
      color = Diff_net_entries_percentage,
      size = Feb_net_entries
    ),
    shape = 16
  ) 
```

```{r}
ggplot(
  data = daily_combined_table_sf %>% filter(
    # net_entries > 10,
    # Feb_net_entries > 10,
    Month == 4,!is.na(Diff_net_entries_percentage),
    DayType == "Weekday"
  ) %>%  mutate(
    Borough = recode(
      Borough,
      Bx = "Bronx",
      Bk = "Brooklyn",
      M = "Manhattan",
      Q = "Queens",
      SI = "Staten Island"
    )
  )
) +
  labs(title = "Average Decline in Daily System Exits (Weekday) between February and April",
       subtitle = "Percentage change, by station",
       fill = "Median Zip Code Income ($)",
       color = "Percentage Change from February",
       size = "Average Daily Exits in February") +
  geom_sf(
    data = zipcode_data_sf,
    aes(fill = MedianIncome, geometry = geometry),
    size = 0.25,
    colour = "gray"
  ) +
  scale_fill_gradient(
    low = "#C9E3F2",
    high = "#071469",
    limits = c(0, 100000),
    oob = scales::squish
  ) +
  scale_color_gradient(
    low = "dark green",
    high = "light green",
    limits = c(-100, -60),
    oob = scales::squish
  ) +
  geom_sf(
    aes(
      geometry = geometry,
      color = Diff_net_exits_percentage,
      size = Feb_net_exits
    ),
    shape = 16
  ) 
```

```{r}
ggplot(data = daily_combined_table_sf %>% filter(
  # net_entries > 10,
  # Feb_net_entries > 10,
  Month == 4,
  !is.na(Diff_net_entries_percentage),
  DayType == "Weekday"
)) + geom_point(aes(x = Diff_net_entries_percentage, y = Diff_net_exits_percentage,
                    color = Borough, size = MedianIncome), shape = 1) +
  labs(title = "Decline in Average Daily Station Entries vs Decline in Average Daily Station Exits",
       subtitle = "From February to April",
       x = "Percentage Decline in Turnstile Entries",
       y = "Percentage Decline in Turnstile Exits",
       size = "Median Income")
```

### Top 10% vs Bottom 10%
```{r}
turnstile_data_grouped_zip <-
  turnstile_data_grouped %>% inner_join(station_complex_to_zip)
income_percentiles <-
  quantile(unique(turnstile_data_grouped_zip$MedianIncome), c(0.2, 0.4, 0.6, 0.8))
turnstile_data_grouped_zip_20 <-
  turnstile_data_grouped_zip %>% filter(MedianIncome <= income_percentiles["20%"])
turnstile_data_grouped_zip_80 <-
  turnstile_data_grouped_zip %>% filter(MedianIncome >= income_percentiles["80%"])
```

```{r}
turnstile_summaries_20 <-
  turnstile_data_grouped_zip_20 %>%
  group_by(datetime = datetime_round) %>%
  summarize(
    entry_sum = sum(net_entries, na.rm = TRUE),
    exit_sum = sum(net_entries, na.rm = TRUE)
  ) %>% ungroup()

turnstile_summaries_80 <-
  turnstile_data_grouped_zip_80 %>%
  group_by(datetime = datetime_round) %>%
  summarize(
    entry_sum = sum(net_entries, na.rm = TRUE),
    exit_sum = sum(net_entries, na.rm = TRUE)
  ) %>% ungroup()
```

```{r}
turnstile_week_percentages_20 <- get_percentage_of_period(turnstile_summaries_20, "week")
turnstile_week_percentages_80 <- get_percentage_of_period(turnstile_summaries_80, "week")
turnstile_month_percentages_20 <- get_percentage_of_period(turnstile_summaries_20, "month")
turnstile_month_percentages_80 <- get_percentage_of_period(turnstile_summaries_80, "month")
```

```{r}
ggplot(data = turnstile_week_percentages_20) +
  geom_point(aes(
    x = as_date("2020-01-01") + (7*turnstile_week_percentages_20$Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour)
  )) +
  geom_line(aes(
    x = as_date("2020-01-01") + (7*turnstile_week_percentages_20$Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour),
    group = as.factor(Hour)
  )) +
  facet_wrap(vars(DayType)) +
  labs(
    title = "Proportion of Daily Total Turnstile Entries by Time Interval (20th Percentile Income)",
    x = "Month",
    y = "Proportion of All Rides",
    color = "Interval Ending At",
    subtitle = paste(
      "Orange Vertical Line is when State of Emergency was Declared in NYC (March 12)",
      "Red Vertical Line is when Statewide Stay-At-Home Order was Enacted (March 22)",
      sep = "\n"
    )
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-12-2020")),
    color = "dark orange",
    linetype = "dashed"
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-22-2020")),
    color = "red",
    linetype = "dashed"
  )
```

```{r}
ggplot(data = turnstile_week_percentages_80) +
  geom_point(aes(
    x = as_date("2020-01-01") + (7*turnstile_week_percentages_80$Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour)
  )) +
  geom_line(aes(
    x = as_date("2020-01-01") + (7*turnstile_week_percentages_80$Period),
    y = ProportionOfDailyEntry,
    color = as.factor(Hour),
    group = as.factor(Hour)
  )) +
  facet_wrap(vars(DayType)) +
  labs(
    title = "Proportion of Daily Total Turnstile Entries by Time Interval (80th Percentile Income)",
    x = "Month",
    y = "Proportion of All Rides",
    color = "Interval Ending At",
    subtitle = paste(
      "Orange Vertical Line is when State of Emergency was Declared in NYC (March 12)",
      "Red Vertical Line is when Statewide Stay-At-Home Order was Enacted (March 22)",
      sep = "\n"
    )
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-12-2020")),
    color = "dark orange",
    linetype = "dashed"
  ) +
  geom_vline(
    mapping = aes(xintercept = mdy("03-22-2020")),
    color = "red",
    linetype = "dashed"
  )
```

```{r}
ggplot(data = turnstile_month_percentages_20 %>% filter(Period %in% c(2, 4)) %>% 
  mutate(Period = month.abb[Period])) +
  geom_point(aes(x = ((Hour - 2) + 24) %% 24,
                 y = ProportionOfDailyEntry,
                 color = as.factor(Period))) +
  geom_line(aes(x = ((Hour - 2) + 24) %% 24,
                y = ProportionOfDailyEntry,
                color = Period,
                group = as.factor(Period))) +
  facet_grid(. ~ DayType) +
  labs(
    x = "Time Of Day",
    y = "Proportion of All Rides",
    title = "Proportion of Daily Total Turnstile Entries by Time Interval in February and April",
    subtitle = "Below 20th Percentile Income\nPoints are at Time Interval Midpoints for Easier Visual Interpretation",
    color = "Month",
    caption = "For instance, 10am in this chart corresponds to 12pm in the previous chart"
  )

ggplot(data = turnstile_month_percentages_80 %>% filter(Period %in% c(2, 4)) %>% 
  mutate(Period = month.abb[Period])) +
  geom_point(aes(x = ((Hour - 2) + 24) %% 24,
                 y = ProportionOfDailyEntry,
                 color = as.factor(Period))) +
  geom_line(aes(x = ((Hour - 2) + 24) %% 24,
                y = ProportionOfDailyEntry,
                color = Period,
                group = as.factor(Period))) +
  facet_grid(. ~ DayType) +
  labs(
    x = "Time Of Day",
    y = "Proportion of All Rides",
    title = "Proportion of Daily Total Turnstile Entries by Time Interval in February and April",
    subtitle = "Above 80th Percentile Income\nPoints are at Time Interval Midpoints for Easier Visual Interpretation",
    color = "Month",
    caption = "For instance, 10am in this chart corresponds to 12pm in the previous chart"
  )
```

